When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-14 paper/paper.tex
Kappa is an ensemble (review 3)

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be
underconstrained, and is likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained,
and is likely to be unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and
problems where $\kappa$ is very small or very large are usually easy. By handling injectivity as a
restriction on the size of the state space rather than as a constraint, we derive
\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left( t^{\underline{p}}
\cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
\right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the sixth
row of \cref{figure:induced}. We see that constrainedness predicts that the central region will
still be relatively difficult for larger pattern graphs: although the problem is overconstrained, it
is less overconstrained than in the regions the Glasgow and LAD solvers found easy.  Thus it seems
that rather than just being a unification of existing generalised heuristic techniques,
constrainedness also gives a better predictor of difficulty than proximity to a phase
transition---our method generates instances where constrainedness and ``close to a phase
transition'' give very different predictions, and constrainedness closely matches the empirical
results.

>>>>

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{DBLP:conf/aaai/GentMPW96}. An ensemble of problem instances with $\kappa < 1$ is
said to be underconstrained, and is likely to be made up of satisfiable instances; an ensemble with
$\kappa > 1$ is overconstrained, and is likely to be made of unsatisfiable instances. Empirically,
problem instances from an ensemble with $\kappa$ close to 1 are hard, and problems drawn from an
ensemble where $\kappa$ is very small or very large are usually easy.

By handling injectivity as a
restriction on the size of the state space rather than as a constraint, we derive
\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left( t^{\underline{p}}
\cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
\right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the sixth
row of \cref{figure:induced}. We see that constrainedness predicts that the central region will
still be relatively difficult for larger pattern graphs: although a family of instances generated
with these parameters is overconstrained, it is less overconstrained than in the regions the Glasgow
and LAD solvers found easy.  Thus it seems that rather than just being a unification of existing
generalised heuristic techniques, constrainedness also gives a better predictor of difficulty than
proximity to a phase transition---our method generates collections of instances where
constrainedness and ``close to a phase transition'' give very different predictions, and
constrainedness closely matches the empirical results.
